Inferring the gravitational potential of the Milky Way with a few precisely measured stars

### 2014-02-07 text/paper2/ms.tex
minor outline revisions

\title{Inferring the gravitational potential of the Milky Way with few precisely measured stars}
\author{Adrian M. Price-Whelan, DWH, KVJ, others?}

@@

[From KVJ:]
- Broader comparison to other methods
	- 'generative methods' more generally (ie N-body, plus Varghese, plus Kuepper plus new possible action-angle models....
	- statistical methods
	- Sanders \& Binney 2013
- Then a short note on worries
	- for generative models
	- for any methods based on action/angle 
- Then how rewinder gets round these????
	- Reference Deg \& Widrow to say they illustrate the idea of the power of multiple streams to back up the bit where you say that's the long-term aim

[Cold streams contain tons of information.]

[GD-1 was fit as if it highlighted an orbit.  That is known to be
  wrong.]

[In the halo, any phase-space structure might be long-lived.  Chemical
  tagging could in principle illuminate it.]

@@

[Sometimes stellar samples might get extremely small.  For example,
  \gaia\ might find some very low-mass cold stellar streams.  For
  another, there might only be a coupld RRL stars in a particular
  stream, but those RRL stars might be individually extremely well
  measured in six-space.]

Even if you have a ton of stars, some might be orders of magnitude more constraining (Bovy, ..., Hogg)

[Obviously a small stellar sample can only tell you a small number of
  things about the Milky Way potential.  The idea is---in the long
  run---to construct likelihood functions for many small stellar
  samples and in the end do inference by multiplying them together.
  Duh!]

[Densities are hard to match -- need a background model, way to go from luminosity to density, needs assumptions about the mass function and structure of the satellite\

@@

[Are there integrable triaxial potentials?]

@@

Hogg:

If interested in oblateness and mass, there are close integrable potentials

- if want to ask the cosmological question (connection to lambda-cdm), need to face non-integrable potentials

- No even approximate way to answer questions about substructure or severe triaxiality that changes with radius

>>>>

\title{Inferring the gravitational potential of the Milky Way with a few precisely measured stars}
\author{Adrian M. Price-Whelan, David W. Hogg, Kathryn V. Johnston, [TODO: order?]}

@@

[TODO: A historical introduction to dark matter, potential measures?]

Major open questions in near-field cosmology: (1) is the Milky Way potential triaxial and if so, how much? (2) How does the potential vary with radius? (3) Is there graininess to the potential? e.g., is there substructure?

Cosmological simulations of galaxy formation in the $\Lambda$CDM paradigm seem to produce triaxial dark matter halos \citep{dubinski91, jing02, kuhlen07, veraciro11}. Halos tend to be more strongly triaxial in dark-matter-only simulations, but still present with the inclusion of baryons. Baryonic processes tend to modify the dark matter structure in the inner galaxy, leading possibly to radially-dependent axis ratios and halo orientation. [statements about integrability]. [if want to connect lambda-CDM to the milky way, we need to face non-integrable potentials]

[something about measuring halo shape with lensing].

[why use tidal streams?] \cite{lm10} used N-body models to ... triaxial.

Another prediction of hierarchical structure formation in $\Lambda$CDM is the presence of substructure.

- Broader comparison to other methods\\
\indent\indent- 'generative methods' more generally (ie N-body, plus Varghese, plus Kuepper plus new possible action-angle models....\\
\indent\indent- statistical methods\\
\indent\indent- Sanders \& Binney 2013\\
- Then a short note on worries\\
\indent\indent- for generative models\\
\indent\indent- for any methods based on action/angle \\
\indent\indent- GD-1 was fit as if it highlighted an orbit.  That is known to be wrong.\\
\indent\indent- Densities are hard to match: need a background model, way to go from luminosity to density, needs assumptions about the mass function and structure of the satellite\\
\indent\indent- Triaxiality / radially-dependent shapes and substructure present problems for integral-based methods\\
- Then how rewinder gets round these????\\
\indent\indent- Reference Deg \& Widrow to say they illustrate the idea of the power of multiple streams to back up the bit where you say that's the long-term aim\\

@@

[Sometimes stellar samples might get extremely small.  For example,
  \gaia\ might find some very low-mass cold stellar streams.  For
  another, there might only be a couple RRL stars in a particular
  stream, but those RRL stars might be individually extremely well
  measured in six-space.]

[Even if you have many stars, some might be orders of magnitude more constraining (Bovy, ..., Hogg)]

@@

@@
